IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-05 paper/formal.tex
Remove invalid case in erasure function.
minor bugs and nits

When defining the security guarantees of an information flow control,
we must characterize what the \emph{secret inputs} of a program are.  Like
other work~\cite{Li+:2010:arrows,Russo+:Haskell08,lio,stefan:addressing-covert},
we specify and prove non-interference using \emph{term erasure}.
%
Intuitively, term erasure allows us to show that an attacker does not learn
any sensitive information from a program if the program behaves identically
(from the attackers point of view) to a program with all sensitive data
``erased''.
%
To interpret a language under information flow control, we define a function |erasef l| that
performs erasures by mapping configurations to erased configurations,
usually by rewriting (parts of) configurations that are more sensitive
than |l| to a new syntactic construct |bullet|.  We can the define
an information flow control language as follows:

@@

\begin{figure} % fig:erasure
\begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|fullconf id il' tS ie| \begin{cases}
    |bullet| & |il' cantFlowTo il| \\
    |fullconf id il' tS ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
    |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
    & \text{the label of thread |id|} \\
    |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|\\
  &|erase il v = v|
\end{align*}
\caption{ Erasure function for tasks, queue maps, message queues, and
configurations.  In all other cases, including target-language constructs,
|erasef il| is applied homomorphically.  \label{fig:erasure} }
\end{figure}

>>>>

When defining the security guarantees of an information flow control,
we must characterize what the \emph{secret inputs} of a program are.  Like
other work~\cite{Li+:2010:arrows,Russo+:Haskell08,lio,stefan:addressing-covert},
we specify and prove non-interference using \emph{term erasure}.
%
Intuitively, term erasure allows us to show that an attacker does not learn
any sensitive information from a program if the program behaves identically
(from the attackers point of view) to a program with all sensitive data
``erased''.
%
To interpret a language under information flow control, we define a function |erasef l| that
performs erasures by mapping configurations to erased configurations,
usually by rewriting (parts of) configurations that are more sensitive
than |l| to a new syntactic construct |bullet|.  We define
an information flow control language as follows:

@@

\begin{figure} % fig:erasure
\begin{align*}
  &|erase il (iconf iS its) =
  iconf (erase il iS) (fltr (\ it . it = bullet) (map (erasef il) its))|\\
  &|erase il (fullconf id il' tS ie) =| \begin{cases}
    |bullet| & |il' cantFlowTo il| \\
    |fullconf id il' tS ie| & \text{otherwise}
  \end{cases} \\
  &|erase il emptyset = emptyset|\\
  &|erase il (bracket iS (mapsto id Q)) =| \begin{cases}
    |erase il iS| & \text{|il' cantFlowTo il|, where |il'| is}\\
    & \text{the label of thread |id|} \\
    |bracket (erase il iS) (mapsto id (erase il Q))| & \text{otherwise}
  \end{cases} \\
  &|erase il Q = filter il Q|
\end{align*}
\caption{ Erasure function for tasks, queue maps, message queues, and
configurations.  In all other cases, including target-language constructs,
|erasef il| is applied homomorphically.  \label{fig:erasure} }
\end{figure}

### 2014-07-05 paper/retrofit.tex
minor bugs and nits

Most modern, dynamic information flow control languages encode
policy by associating a \emph{label}
with data, which encodes whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order |canFlowTo|, join |lub|, and meet |glb|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

As opposed to more traditional fine-grained language-based IFC systems that
associate labels with values~\cite{sabelfeld2003language,
Austin:Flanagan:PLAS09}, we follow~\cite{Zeldovich:2006, lio} in taking a
coarse-grained floating-label approach and associating labels with
\emph{tasks}.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.
%
Hence before performing a reads or a write, the IFC monitor inspects the
current label to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|).
    \item Operations for inspecting (\textbf{getLabel}) and modifying
    (\textbf{setLabel}) the current label of the task (a task can only increase its label),
    \item Operations for non-blocking inter-task communication (\textbf{send}
    and \textbf{recv}), which interact with the global store of per-task
    message queues |iS|, and
%%     Labels are associated with message at the time of
%%     the send; the label on a message must always be at least as sensitive as that
%%     of the sending thread, since, conceptually, once a message is put on a
%%     message queue it is no longer protected by the current label, rather it is
%%     protected by the message label.
%%     Dually, when reading a message with |recv| the current label of the
%%     receiver must be at least as high as the message label.
    \item A sandboxing operation used to spawn new isolated tasks. In
    concurrent settings |fork| corresponds to a fork-like primitive,
    whereas in a
    sequential setting, it more closely resembles the LIO's
    |toLabeled|~\cite{lio}, and the Breeze
    bracket~\cite{Hritcu:2013:YIB:2497621.2498098}.
\end{itemize}

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.~\cite{Matthews:2007:OSM:1190216.1190220}
    %
In the target language, an IFC expression appears as |TI ie| (``target-outside,
IFC-inside''); in the IFC language, a target language expression appears as |IT
te| ( ``IFC-outside, target-inside'').
    \item We reinterpret |CE| to be evaluation contexts on task lists, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task list, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task lists.  The correspondence is simple: a task executes a step and then is rescheduled in the task list according to schedule policy |alpha|.
    Figure~\ref{fig:scheduler} defines two concrete schedulers.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    and intermediating between the borders of the two languages.
\end{itemize}

@@

Rules \textsc{I-done} and \textsc{I-noStep} define the behavior of the system
when the current thread has reduced to a value, or gotten stuck, respectively.
%
While these definitions simply rely on the underlying scheduling policy,
|alpha|, to modify the task list, as we describe in Sections~\ref{sec:formal}
and~\ref{sec:extensions}, these rules (notably, \textsc{I-noStep}) are crucial
to proving our security guarantees.
%
For instance, it is unsafe for the whole system to get stuck if a particular
task gets stuck, since a sensitive thread may then leverage this to leak
information through the termination channel.
%
Instead, as our example |roundrobinf| scheduler shows, such tasks should simply
be removed from the task list.

@@

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.
\begin{align*}
    & |let i = TI (sandbox (brecvf x i' (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}
It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the sandboxed task, but it will be a dangling pointer; the
task will get stuck (under the |roundrobinf| scheduler the \textsc{I-noStep} will in turn remove it from the task list).  Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.

@@

Our general embedding approach provides these semantics by construction.
%
To show this, we first define an encoding of the sequential |toLabeled|
construct in our framework:
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId; | \\
    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
When using the sequential scheduler |seqf|, the sandboxed task runs to
completion before returning to the task that sandboxed it. 
(Here, a task list can be interpreted as a number of information flow aware
stack frames.)
To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv}
rule.\footnote{A keen reader may note that with a single, per-task channel, we
might mix up our messages.  Indeed, a proper development requires the ability
to allocate fresh channels; for purposes of simplicity, this has been omitted.}

Now suppose that our |targetLangML| is extended with a single exception
expression |fail|, which reduces to |fail|, removing all containing |tE|
evaluation contexts, i.e., it propagates up to a |IT (-)| boundary.
%
In this setting, it is clear that exceptions cannot cross boundaries.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
%
Just like in both LIO and Breeze, however, only when calling |recv| on such messages
will the exception be propagated.
%
This of course, is not a leak since the current label must be raised before
receiving sensitive message, regardless of their content.

>>>>

Most modern, dynamic information flow control languages encode
policy by associating a \emph{label}
with data, which encodes whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order $\sqsubseteq$, join |lub|, and meet |glb|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

As opposed to more traditional fine-grained language-based IFC systems that
associate labels with values~\cite{sabelfeld2003language,
Austin:Flanagan:PLAS09}, we follow~\cite{Zeldovich:2006, lio} in taking a
coarse-grained floating-label approach and associating labels with
\emph{tasks}.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.
%
Hence before performing a read or a write, the IFC monitor inspects the
current label to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|).
    \item Operations for inspecting (\textbf{getLabel}) and modifying
    (\textbf{setLabel}) the current label of the task (a task can only increase its label).
    \item Operations for non-blocking inter-task communication (\textbf{send}
    and \textbf{recv}), which interact with the global store of per-task
    message queues |iS|.
%%     Labels are associated with message at the time of
%%     the send; the label on a message must always be at least as sensitive as that
%%     of the sending thread, since, conceptually, once a message is put on a
%%     message queue it is no longer protected by the current label, rather it is
%%     protected by the message label.
%%     Dually, when reading a message with |recv| the current label of the
%%     receiver must be at least as high as the message label.
    \item A sandboxing operation used to spawn new isolated tasks. In
    concurrent settings |fork| corresponds to a fork-like primitive,
    whereas in a
    sequential setting, it more closely resembles the LIO
    |toLabeled|~\cite{lio} and Breeze
    bracket~\cite{Hritcu:2013:YIB:2497621.2498098} constructs.
\end{itemize}

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of
      both languages to allow for embeddings terms in one language to
      be embedded in the other, as
      in~\cite{Matthews:2007:OSM:1190216.1190220}.
    %
In the target language, an IFC expression appears as |TI ie| (``target-outside,
IFC-inside''); in the IFC language, a target language expression appears as |IT
te| ( ``IFC-outside, target-inside'').
    \item We reinterpret |CE| to be evaluation contexts on task lists, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task list, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task lists.  The correspondence is simple: a task executes a step and then is rescheduled in the task list according to schedule policy |alpha|.
    Figure~\ref{fig:scheduler} defines two concrete schedulers.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    and intermediating between the borders of the two languages.
\end{itemize}

@@

Rules \textsc{I-done} and \textsc{I-noStep} define the behavior of the system
when the current thread has reduced to a value, or gotten stuck, respectively.
%
While these definitions simply rely on the underlying scheduling policy,
|alpha|, to modify the task list, as we describe in Sections~\ref{sec:formal}
and~\ref{sec:extensions}, these rules (notably, \textsc{I-noStep}) are crucial
to proving our security guarantees.
%
For instance, it is unsafe for the whole system to get stuck if a particular
task gets stuck, since a sensitive thread may then leverage this to leak
information through the termination channel.
%
Instead, as our example round-robin (|roundrobinf|) scheduler shows,
such tasks should simply be removed from the task list.

@@

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}
It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the sandboxed task, but it will be a dangling pointer; the
task will get stuck (under the |roundrobinf| scheduler the \textsc{I-noStep} will in turn remove it from the task list).  Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.

@@

Our general embedding approach provides these semantics by construction.
%
To show this, we first define an encoding of the sequential |toLabeled|
construct in our framework:
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId in| \\
%%    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab let j = TI (sandbox (IT (TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
When using the sequential scheduler |seqf|, the sandboxed task runs to
completion before returning to the task that sandboxed it. 
(Here, a task list can be interpreted as a number of information flow aware
stack frames.)
To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv}
rule.\footnote{A keen reader may note that with a single, per-task channel, we
might mix up our messages.  Indeed, a proper development requires the ability
to allocate fresh channels; for purposes of simplicity, this has been omitted.}

Now suppose that our |targetLangML| is extended with a single exception
expression |fail|, which reduces to |fail|, removing all containing |tE|
evaluation contexts, i.e., it propagates up to a |IT (-)| boundary.
%
In this setting, it is clear that exceptions cannot cross boundaries.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
%
Just like in LIO and Breeze, however, only when calling |recv| on such messages
will the exception be propagated.
%
This of course, is not a leak since the current label must be raised before
receiving sensitive message, regardless of their content.
\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}
